
# This is the assignment submission for Reproducible Research Week2

## Loading the data file into a data frame (Data file is already downloaded and extracted)
```{r}
activity <- read.csv("activity.csv", as.is = TRUE)
```

##Remove the NA values and store in a separate structure for future use
```{r}
clean<- activity[complete.cases(activity), ]
```

## Calculate the total number of steps taken per day
```{r}
steps_per_day <- aggregate(steps ~ date, clean, sum)
```

## Create a histogram of no of steps per day
```{r}
hist(steps_per_day$steps, main = "Histogram of total number of steps per day", xlab = "Steps per day", col = "Green")
```

## Calculate the mean and median of the total number of steps taken per day
```{r}
mean(steps_per_day$steps)
median(steps_per_day$steps)
```

## Calculate average steps per interval for all days 
```{r}
avg_steps_per_interval <- aggregate(steps ~ interval, clean, mean)
```

## Calculate average steps per day for all intervals - Not required, but for my own sake 
```{r}
avg_steps_per_day <- aggregate(steps ~ date, clean, mean)
```

## Plot the time series with appropriate labels and heading
```{r}
plot(avg_steps_per_interval$interval, avg_steps_per_interval$steps, type='l', col=1, main="Average number of steps by Interval", xlab="Time Intervals", ylab="Average number of steps")
```

## Identify the interval index which has the highest average steps
```{r}
interval_idx <- which.max(avg_steps_per_interval$steps)
```

## Identifing the specific interval and the average steps for that interval
```
print (paste("The interval with the highest avg steps is ", avg_steps_per_interval[interval_idx, ]$interval, " and the no of steps for that interval is ", round(avg_steps_per_interval[interval_idx, ]$steps, digits = 1)))
```

## Calculatinge number of rows with missing values
```
missing_value_act <- activity[!complete.cases(activity), ]
nrow(missing_value_act)
```

## Loop thru all the rows of activity, find the one with NA for steps.
## For each identify the interval for that row
## Then identify the avg steps for that interval in avg_steps_per_interval
## Substitute the NA value with that value
```{r}
for (i in 1:nrow(activity)) {
  if(is.na(activity$steps[i])) {
    val <- avg_steps_per_interval$steps[which(avg_steps_per_interval$interval == activity$interval[i])]
    activity$steps[i] <- val 
  }
}
```

## Aggregate the steps per day with the imputed values
```{r}
steps_per_day_impute <- aggregate(steps ~ date, activity, sum)
```

## Draw a histogram of the value 
```{r}
hist(steps_per_day_impute$steps, main = "Histogram of total number of steps per day (IMPUTED)", xlab = "Steps per day")
```

## Compute the mean and median of the imputed value
## Calculate the mean and median of the total number of steps taken per day
```{r}
round(mean(steps_per_day_impute$steps))
median(steps_per_day_impute$steps)
```

## weekday Pattern
```{r}
week_day <- function(date_val) {
  wd <- weekdays(as.Date(date_val, '%Y-%m-%d'))
  if  (!(wd == 'Saturday' || wd == 'Sunday')) {
    x <- 'Weekday'
  } else {
    x <- 'Weekend'
  }
  x
}
```

## Apply the week_day function and add a new column to activity dataset
```{r}
activity$day_type <- as.factor(sapply(activity$date, week_day))
```

##load the ggplot library
```{r}
library(ggplot2)
```

## Create the aggregated data frame by intervals and day_type
```{r}
steps_per_day_impute <- aggregate(steps ~ interval+day_type, activity, mean)
```

## Create the plot
```{r}
plt <- ggplot(steps_per_day_impute, aes(interval, steps)) +
  geom_line(stat = "identity", aes(colour = day_type)) +
  theme_gray() +
  facet_grid(day_type ~ ., scales="fixed", space="fixed") +
  labs(x="Interval", y=expression("No of Steps")) +
  ggtitle("No of steps Per Interval by day type")
print(plt)
```


